Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2023-11-21 ATACCC.tex
fix maths formatting

Viral load exhibits exponential growth followed by exponential decline.
\Textcite{hakkiOnset} use a model which is approximately piecewise linear but continuous.
A continuous viral load trajectory should give better performance when using gradient-based inference methods such as Hamiltonian Monte Carlo (HMC)\todo{cite paper or ref section}.
The model for the viral load at time $t$, $\hat{v}_i(t)$, in individual $i$ is shown in \cref{ATACCC:fig:viral-load-model} and given by:
\begin{align}
v_i(\tau) = v_{\max,i} \frac{a_i+b_i}{b_ie^{-a_i(\tau-\tau_{\max,i})} + a_ie^{b_i(\tau-\tau_{max,i})}}. \label{ATACCC:eq:viral-load}
\end{align}
Time is continuous and relative to the time that individual $i$ was first tested.
$v_{\max,i}$ is the maximum viral load, which occurs at time $\tau_{\max,i}$.
As $|\tau - tau_\max|$ becomes large, the viral load trajectory tends towards exponential growth (before the peak) or decline (after the peak) in viral load at rates $a_i$ and $b_i$ respectively.
Before the peak, this is because $b(\tau-\tau_{\max,i})$ is small and hence $a\exp(b(\tau-\tau_{\max,i}))$ is negligible.
After the peak, this is because $b\exp(-a(\tau-\tau_{\max,i}))$ is negligible.
This happens quickly for realistic values of the parameters.
Therefore, the log viral load is approximately piecewise linear (see \cref{ATACCC:fig:viral-load-model} for a representative viral load trajectory).
$\vec{\phi}_i = (v_{\max,i}, a_i, b_i)^T$ are the random effects for individual $i$.
$\tau_{\max,i}$ is inferred independently for each individual because the time from first test to peak is not a biological concept shared between individuals.
\begin{figure}
  \centering \includegraphics{ATACCC/typical_trajectory}
  \caption[Example viral load model trajectory.]{Example viral load trajectory produced by \textcite{hakkiOnset}'s viral load model, using realistic parameter values. The increase in log viral load is approximately linear with a gradient of $a_i$, here 5. The decrease is also linear, with gradient $-b_i$, here -1.4. The peak is at time $\tau_{\max,i}$, here 0, and log viral load $\log v_{\max,i}$, here = 15. The smoothing (lack of linearity) near the peak is visible. \label{ATACCC:fig:viral-load-model}}
\end{figure}

>>>>

Viral load exhibits exponential growth followed by exponential decline.
\Textcite{hakkiOnset} use a model which is approximately piecewise linear but continuous.
A continuous viral load trajectory should give better performance when using gradient-based inference methods such as Hamiltonian Monte Carlo (HMC)\todo{cite paper or ref section}.
The model for the viral load at time $t$, $\hat{v}_i(t)$, in individual $i$ is shown in \cref{ATACCC:fig:viral-load-model} and given by:
\begin{align}
v_i(\tau) = v_{\max,i} \frac{a_i+b_i}{b_ie^{-a_i(\tau-\tau_{\max,i})} + a_ie^{b_i(\tau-\tau_{max,i})}}. \label{ATACCC:eq:viral-load}
\end{align}
Time is continuous and relative to the time that individual $i$ was first tested.
$v_{\max,i}$ is the maximum viral load, which occurs at time $\tau_{\max,i}$.
As $\lvert \tau - \tau_{\max} \rvert$ becomes large, the viral load trajectory tends towards exponential growth (before the peak) or decline (after the peak) in viral load at rates $a_i$ and $b_i$ respectively.
Before the peak, this is because $b(\tau-\tau_{\max,i})$ is small and hence $a\exp(b(\tau-\tau_{\max,i}))$ is negligible.
After the peak, this is because $b\exp(-a(\tau-\tau_{\max,i}))$ is negligible.
This happens quickly for realistic values of the parameters.
Therefore, the log viral load is approximately piecewise linear (see \cref{ATACCC:fig:viral-load-model} for a representative viral load trajectory).
$\vec{\phi}_i = (v_{\max,i}, a_i, b_i)^T$ are the random effects for individual $i$.
$\tau_{\max,i}$ is inferred independently for each individual because the time from first test to peak is not a biological concept shared between individuals.
\begin{figure}
  \centering \includegraphics{ATACCC/typical_trajectory}
  \caption[Example viral load model trajectory.]{Example viral load trajectory produced by \textcite{hakkiOnset}'s viral load model, using realistic parameter values. The increase in log viral load is approximately linear with a gradient of $a_i$, here 5. The decrease is also linear, with gradient $-b_i$, here -1.4. The peak is at time $\tau_{\max,i}$, here 0, and log viral load $\log v_{\max,i}$, here = 15. The smoothing (lack of linearity) near the peak is visible. \label{ATACCC:fig:viral-load-model}}
\end{figure}
